---
title: "Calibrating Jomon Ceramic Chronology into Calendar Dates"
author: "E.Crema"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a short tutorial for using the posterior samples and the methodology presented in [Crema and Kobyashi 2020](https://doi.org/10.1016/j.jas.2020.105136) to estimate the absolute calendar dates of Jomon ceramic phases. 

## Setup

In order calibrate ceramic phases you will need R statistical computing language installed in your machine and have the following three files in your working directory: [posteriorSamples.RData](https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R_images/posteriorSamples.RData),  [utilities.R](https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R/mcsim.R), and [mcsim.R](https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R/mcsim.R). 


```{r,eval=FALSE}
source('mcsim.R')
load('posteriorSamples.RData')
```

Or alternativately you can download and load the R image containing the posteriors and `mcsim()` function directly using the following command:
```{r}
#source('https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R/utilities.R')
#source('https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R/mcsim.R')
#load(file = url('https://github.com/ercrema/jomonPhasesAndPopulation/raw/master/R_images/posteriorSamples.RData'))
source('https://github.com/ercrema/jomonPhasesAndPopulation/raw/intcal20version/R/mcsim.R')
source('https://github.com/ercrema/jomonPhasesAndPopulation/raw/intcal20version/R/utilities.R')
load(file = url('https://github.com/ercrema/jomonPhasesAndPopulation/raw/intcal20version/R_images/posteriorSamples.RData'))
```

## Data Preparation

The main argument for the `mcsim()` is a data.frame where each row contains an object to be dated with two columns (with names `$StartPhase` and `$EndPhase`) containing the start and the end Jomon ceramic phases for the timespan of possible existence of each event to be dated. The name of the ceramic phases should follow the convention used in Table 1 of [Crema and Kobyashi 2020](https://doi.org/10.1016/j.jas.2020.105136) reported in a redacted form below:

<!-- ************************************************************************** -->
<table cellspacing="0" border="0">
	<colgroup width="217"></colgroup>
	<colgroup width="179"></colgroup>
	<colgroup width="137"></colgroup>
	<colgroup width="349"></colgroup>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" height="20" align="center" valign=middle><b><font size=3 color="#000000">Period</font></b></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><b><font size=3 color="#000000">Ceramic Phase Name</font></b></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><b><font size=3 color="#000000">Kobayashi 2017</font></b></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><b><font size=3 color="#000000">Ceramic Phase in Suzuki's Pithouse Data</font></b></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" height="20" align="center" valign=bottom><font size=3 color="#000000">End of Upper Palaeolithic</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S0</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S0</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Mumon doki</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=4 height="178" align="center" valign=middle><font size=3 color="#000000">Incipient Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S1.1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S1-1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000"><br>Ryūkisenmon-kei<br></font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S1.2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S1-2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Biryūkisenmon-kei; Tsumegatamon-kei</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S2.1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S2-1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000"><br>Tsumegatamon-kei; Ōnatsumon-kei<br></font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S2.2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S2-2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000"><br>Tajōmon-kei<br></font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=9 height="196" align="center" valign=middle><font size=3 color="#000000">Initial Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=4 align="center" valign=middle><font face="Calibri" size=3 color="#000000">S3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">S3-1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=4 align="left" valign=middle><font size=3 color="#000000">Igusa I; Igusa II; Daimaru; Natsushima; Inaridai; Tateno; Inarihara; Ōurayama; Hanawadai 1; Hanawadai 2; Hirasaka;</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">S3-2</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">S3-3</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">S3-4</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Mito; Lower Tado; Upper Tado; Hosokubo</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Shiboguchi; Nojima</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ugajimadai</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Lower Kayama; Upper Kayama</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S8</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">S8</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Uenoyama; Irumi I; Irumi II ; Ishiyama; Okkōshi; Tenjinyama; Kaminokidai; Shioya; Shimoyoshii</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=7 height="173" align="center" valign=middle><font size=3 color="#000000">Early Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Lower Hanazumi</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Sekiyama; Futatsuki; Kaminoki</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kurohama; Ario</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000"><br>Moroiso a; Minamiōhara<br></font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Moroiso b; Uehara</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Moroiso c; Hinata I; Kagobatake I; Shitajima</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">Z7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Jūsanbodai; Hinata II; Kagohata II</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=18 height="370" align="center" valign=middle><font size=3 color="#000000">Middle Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Goryōgadai 1</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C234</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C2~C4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Goryōgadai 2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C56</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="left" valign=middle><font size=3 color="#000000">Atamadai 1a; Atamadai 1b; Mujinasawa; Katsuzaka I; Aramichi</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C6</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C78</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="left" valign=middle><font size=3 color="#000000">Atamadai 2; Atamadai 3; Katsuzaka 2; Tōnai I; Tōnai II;</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C8</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C9</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C9a</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="left" valign=middle><font size=3 color="#000000">Idojiri I;Idojiri III; Katsuzaka 3; Atamadai 4</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C9bc</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C10</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C10a</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="left" valign=middle><font size=3 color="#000000">Daigi 8a; Kasori E1 (EI)* Sori I</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C10b</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C10c</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C11</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C11ab</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=2 align="left" valign=middle><font size=3 color="#000000">Daigi 8b; Kasori E2 (EI)*; Sori II; Sori III</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C11c</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="center" valign=middle><font face="Calibri" size=3 color="#000000">C12</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C12a</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="left" valign=middle><font size=3 color="#000000">Daigi 9; Kasori E3 (EII-EIII)*; Sori IV;SoriV</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C12b</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">C12c</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C13</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C13</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kasori E4 (EIV)*; Daigi 10;</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">C14</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000"> -</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kasori EV; Daigi 10;</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=10 height="201" align="center" valign=middle><font size=3 color="#000000">Late Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="center" valign=middle><font face="Calibri" size=3 color="#000000">K1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">K1-1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=3 align="left" valign=middle><font size=3 color="#000000">Shōmyōji 1;Shōmyōji 2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">K1-2</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">K1-3</font></td>
		</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Horinouchi 1</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Horinouchi 2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kasori B1</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kasori B2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Kasori B3</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K7</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Takaihigashi; Sōya</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">K8</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font size=3 color="#000000">K8</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Angyō 1; Angyō 2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" rowspan=6 height="120" align="center" valign=middle><font size=3 color="#000000">Final Jomon</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B1</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora B; Angyō 3a</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B2</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora BC; Angyō 3b</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B3</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora C1; Angyō 3c; Maeura 1</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B4</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora C2; Angyō 3d; Maeura 2</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B5</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora A; Chiami; Kōri I</font></td>
	</tr>
	<tr>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="center" valign=middle><font face="Calibri" size=3 color="#000000">B6</font></td>
		<td style="border-top: 1px solid #000000; border-bottom: 1px solid #000000; border-left: 1px solid #000000; border-right: 1px solid #000000" align="left" valign=middle><font size=3 color="#000000">Ōbora A'; Arami; Kōri II</font></td>
	</tr>
</table>
<!-- ************************************************************************** -->

For example, suppose you have four pit-dwellings dated as follows: 

* PithouseA; Daigi8a ~ Daigi8b
* PithouseB; Horinouchi ~ Kasori B
* PithouseC; Horinouchi ~ Kasori B
* PithouseD; Moroiso a

The intput data.frame would be constructed as follows:

```{r}
StartPhase = c('C10','K2','K2','Z4')
EndPhase = c('C11','K6','K6','Z4')
df = data.frame(StartPhase=StartPhase,EndPhase=EndPhase)
rownames(df) = c('A','B','C','D') #this is not strictly necessary 
print(df)
```

Notice that there are as many rows as the number of entities to be dated (i.e. chronologically identical iterms should be repeated) and that entities with no _Phase assignment uncertainty_ (i.e. assigned to a single phase) would have identical entries for the `$StartPhase` and `$EndPhase` fields.

## Using the mcsim() function

The `mcsim()` function randomly generates user-defined number (argument `nsim`) of calendar dates taking into account _phase assignment uncertainty_, _within-phase uncertainty_, and _phase boundary uncertainty_. The _within-phase uncertainty_  and _phase boundary uncertainty_ are modelled by providing pre-processed posterior samples contained the image file `posteriorSamples.RData`. These includes trapezoid (`postTrapezoid`), Gaussian (`postGaussian`), and uniform (`postUniform`) distributions. _Phase assignment uncertainty_ is controlled by the argument `weights` and could be one between `equal` (for equal probability assigned to each possible phase in the interval),  `variance` (for probabilities proportional to the duration of possibles phases in the interval), and `custom` (for user defined probabilities). For example, the following generates 1000 calendar dates using the Trapezoid distribution and a duration based modelling of _phase assignment uncertainty_:

```{r}
res=mcsim(df,nsim=1000,posterior=postTrapezoid,weights="variance")
```

The resulting object is a matrix with each row corresponding to each event. Thus the estimated age of pithouse A would be:

```{r}
eventA = res[1,]
plot(density(eventA),xlim=c(5200,4500),xlab='Cal BP',main='Event A')
abline(v=median(eventA),lwd=2,lty=2)
quantile(eventA,prob=c(0.025,0.975)) #95% percentile interval
```

The 'custom' option for the weights argument allows for user-defined probabilities to be assigned for each phases within the interval of possible existence for each object. This is defined by assigning a list to the argument `weightList`. For example:

```{r}
w = vector('list',length=4)
w[[1]] = c(0.3,0.7)
w[[2]] = c(0.2,0.3,0.2,0.2,0.1)
w[[3]] = c(0.1,0.3,0.3,0.2,0.1)
w[[4]] = c(1)

res=mcsim(df,nsim=1000,posterior=postTrapezoid,weights="variance",weightList = w)
```

It is also possible to provide for each element of `weightList` a matrix instead of vector, with the number of columns corresponding to the number of phases within the timespan. The `mcsim()` function will randomly select a row from sum matrix for each iteration. This approach can, for example, be used when the _phase assignement uncertainty_ is modelled based on the observed frequencies of events for each phase. The script below illustrates a workflow where the Dirichlet distribution is used as a conjugate prior of a multinomial distribution:

```{r}
### Function for generating sample probabilities from the dirichlet distribution ### 
rdirichlet<-function (n, alpha) 
{
  l <- length(alpha)
  x <- matrix(rgamma(l * n, alpha), ncol = l, byrow = TRUE)
  sm <- x %*% rep(1, l)
  x/as.vector(sm)
}


# Hypothetical Observed Frequencies of Events 
Z4 = 10
C10 = 67
C11 = 23
K2 = 1
K3 = 2
K4 = 34
K5 = 67
K6 = 23

#
w2 = vector('list', length=4)
w2[[1]]=rdirichlet(n=1000,alpha=c(C10,C11))
w2[[2]] = w2[[3]]=rdirichlet(n=1000,alpha=c(K2,K3,K4,K5,K6))
w2[[4]] = c(1)

res=mcsim(df,nsim=1000,posterior=postTrapezoid,weights="custom",weightList = w2)
```


